\clearpage
\begin{landscape}
\section{Tracability}

\centering
\noindent\begin{longtable*}{|c|m{10cm}|c|c|m{4cm}|}
\hline
\textbf{Requirement ID} &
\multicolumn{1}{c|}{\textbf{Requirement Title}} &
\textbf{Full Requirement} &
\textbf{Addressed by Design} &
\multicolumn{1}{c|}{\textbf{Addressed by Test}} \\ \hline

Req:ASE &
The framework shall allow researchers to provide search engines for the system to use. &
Page \pageref{sec:reqase} &
Section \ref{sec:mulsearchen} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}}\\ \hline

Req:ASM &
The framework shall allow researchers to provide suitability measures for the system to use. &
Page \pageref{sec:reqasm} &
Section \ref{sec:mulsuitmeas} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}}\\ \hline

Req:QAO &
The solution of a search, a quantum algorithm, shall be presented to the user as a list of instructions. &
Page \pageref{sec:reqqao} &
Section \ref{sec:quantalgs} &
\multicolumn{1}{c|}{Section \ref{sec:algtests}} \\ \hline

Req:CV &
The system shall provide visualisation of the circuit produced by the solution of the search for a system of a user specified number of qubits. &
Page \pageref{sec:reqcv} &
Section \ref{sec:quantumcircuits} &
\multicolumn{1}{c|}{Section \ref{sec:circtests}} \\ \hline

Req:TPS &
The framework shall be able to be embedded in third party software. &
Page \pageref{sec:reqtps} &
Section \ref{sec:thirdpartyblackbox} &
Code Review of Client to ensure only interface knowledge is required\\ \hline

Req:DST &
The framework shall provide a standardised definition format for users to specify the target of the search. &
Page \pageref{sec:reqdst} &
Section \ref{sec:testsuitestruc} &
\multicolumn{1}{c|}{Section \ref{sec:testsuitetests}} \\ \hline

Req:UCF &
The customisation of the framework shall be provided through a series of configuration files. &
Page \pageref{sec:requcf} &
Section \ref{sec:manclasses} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}} \\ \hline

Req:PGAI &
The framework shall provide implementations of all gates specified in Figure \ref{fig:providedgates}. The framework shall provide algorithm instructions for each of these gates and for the instantiation of the Controlled-U gate with all single qubit gates. &
Page \pageref{sec:reqpgai} &
Sections \ref{sec:quantalgs}, \ref{sec:quantumgates} &
\multicolumn{1}{c|}{Sections \ref{sec:gatetests}, \ref{sec:algtests}} \\ \hline

Req:ACS &
The system shall provide the iterate control structure and support nested iterate instructions. &
Page \pageref{sec:reqacs} &
Section \ref{sec:quantalgs} &
\multicolumn{1}{c|}{Sections \ref{sec:gatetests}, \ref{sec:algtests}} \\ \hline

Req:PCA &
The framework shall be able to produce a circuit, for any given number of qubits, from a quantum algorithm. &
Page \pageref{sec:reqpca} &
Sections \ref{sec:quantumcircuits}, \ref{sec:quantumgates} &
Code review of both the circuit implementation and all of the gate implementations to ensure the logic does not rely on upper bound to the system size.
The only restriction that is present is where $2^{System Size}>Integer.MAX\_VALUE$. \\ \hline

Req:CS &
The framework shall provide the simulation of a circuit given an initial state. &
Page \pageref{sec:reqcs} &
Section \ref{sec:quantumgates} &
\multicolumn{1}{c|}{Sections \ref{sec:gatetests}, \ref{sec:cirevaltests}} \\ \hline

Req:SBSSE &
The framework shall provide a way to perform step-by-step evaluation of a circuit given an initial state. &
Page \pageref{sec:reqsbsse} &
Section \ref{sec:sbsecireval} &
\multicolumn{1}{c|}{Section \ref{sec:cirevaltests}} \\ \hline

&&&&
\\ \hline

Req:SSE &
The tool shall provide at least one implemented search engine. &
Page \pageref{sec:reqsse} &
Section \ref{sec:provsearcheng} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}} \\ \hline

Req:SSM &
The tool shall provide at least one implemented suitability measure. &
Page \pageref{sec:reqssm} &
Section \ref{sec:provsuitmeas} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}} \\ \hline

Req:SST &
The tool shall provide a number of search targets with known outputs. &
Page \pageref{sec:reqsst} &
Section \ref{sec:provsearchprobs} &
\multicolumn{1}{c|}{Section \ref{sec:manclasstests}} \\ \hline

Req:SES &
The GUI shall provide a user with a selection of search engines to use in a search. &
Page \pageref{sec:reqses} &
Section \ref{sec:seselecdes} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:SMS &
The GUI shall provide a user with a selection of suitability measures to use in a search. &
Page \pageref{sec:reqsms} &
Section \ref{sec:smselecdes} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:STS &
The GUI shall provide a user with a selection of search targets to be used as the search goal. &
Page \pageref{sec:reqsts} &
Section \ref{sec:spselecdes} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:STC &
The GUI shall provide a way for users to create a new search target without needing to explicitly write a configuration file. &
Page \pageref{sec:reqstc} &
Section \ref{sec:guisearchcreateed} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:STE &
The GUI shall provide a way for users to edit the contents of a previously created search target without manual editing of the configuration file. &
Page \pageref{sec:reqste} &
Sections \ref{sec:indtestsuiteeditor}, \ref{sec:guisearchcreateed} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:LSTPDC &
The GUI shall provide a way to import a predefined search target from a configuration file. &
Page \pageref{sec:reqlstpdc} &
Section \ref{sec:guisearchcreateed} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:SV &
The GUI shall provide a way to visualise any quantum state. &
Page \pageref{sec:reqsv} &
Section \ref{sec:repres} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:RSR &
The GUI shall provide a way to report the search result, a quantum algorithm, to the user. &
Page \pageref{sec:reqrsr} &
Section \ref{sec:repres} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:GCV &
Given a quantum algorithm and a system size, the GUI shall produce a visualisation of the resulting circuit. &
Page \pageref{sec:reqgcv} &
Section \ref{sec:repres} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:GSBSSE &
The GUI shall provide a way to perform, control and visualise the step-by-step state evolution for an intial state and circuit. &
Page \pageref{sec:reqgsbsse} &
Section \ref{sec:repres} &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:TT &
The GUI shall provide user help through the use of tooltips. &
Page \pageref{sec:reqtt} &
TODO &
\multicolumn{1}{c|}{Section \ref{sec:clientguitests}} \\ \hline

Req:POR &
The framework, fully implemented tool and the GUI shall be able to be used on a range of Operating Systems. &
Page \pageref{sec:reqpor} &
TODO &
\multicolumn{1}{c|}{Section \ref{sec:systests}} \\ \hline

Req:USE &
Using either the fully implemented tool or the GUI a user shall be able to start a search within 30 seconds. &
Page \pageref{sec:requse} &
TODO &
\multicolumn{1}{c|}{Section \ref{sec:systests}} \\ \hline

\end{longtable*}
\end{landscape}
\clearpage