\chapter{Requirements}
\label{sec:reqs}

This project aims to provide software to facilitate research into quantum computation through heuristic search.
As an initial architectural decision the software is broken down into three sections.

The first section is a plug-and-play research framework that will enable the search for quantum algorithms by providing developed peripheral utilities such as circuit construction and simulation.
The framework will allow researchers to perform research into search techniques and suitability measures without reimplementing peripheral, non-research, tasks.
The framework is not intended as a standalone application but as a library that can be incorporated into other software.

The second section is a prototype implementation of at least one search technique and suitability measure
The inclusion of these modules shall be used to indicate how the framework is intended to be used by researchers and demonstrate the plug-and-play ability of the framework.

The third section is a client with a graphical user interface for the framework.
A client is necessary to allow the toolkit to be used as a standalone application.
This may seem as though as it should be included as part of the framework, however the distinction is made to highlight the different aims.
The framework is intended to provide the structure and utilities, such as circuit simulation, that can be included as a single module in any application, for example a general search application or a quantum IDE.
Some researchers may not wish to develop an application in which to embed the framework.
The client provides a basic user interface to the framework for such researchers.

% The requirements were maintained using an online tool called ReqMan by RequirementOne\cite{reqman}.

% \section{Purpose}
% \subsection{Framework}
% The framework is aimed to facilitate research into quantum computation through heuristic search.
% 
% \subsection{Implemented Search Prototype}
% The implemented prototype is to provide 
% 
%  of a Quantum Algorithm search tool using the framework is to provide a working toolkit for researchers interested in finding Quantum Algorithms rather than the search techniques to find Quantum Algorithms.
% As the toolkit will use the framework it will also provide a potential foundation for future research into the search techniques.
% 
% \subsection{Client GUI}
% The client GUI will provide an interface that should make the toolkit more accessible for researchers.
% Without the GUI provided, researchers would have to either embed the toolkit in their own application or within their own specific GUI.
% This is likely to reduce the potential use of the toolkit in the academic community.
% One of the main focusses of the toolkit is to try and provide a standardised framework for research of Quantum Algorithms.
% Not providing a GUI, resulting in many bespoke GUIs, goes against this focus.
% This is not to say that inclusion of the framework in 3rd party systems or improvement to the GUI is not encouraged.

% \section{Scope}
% The requirements listed in the following section cover the framework and the client GUI.
% The full implementation, based on the QPace IV suite, is not covered by these requirements.
% This is to improve the clarity between the framework and the implementation.
% 
% The requirements of the full implementation are listed in a separate section following the framework and client GUI requirements.

\section{Definition, Acronyms, and Abbreviations}
\label{sec:reqdefs}
The definitions given here are consistent with those used in the rest of the document but are summarised here for convenience.

\textbf{System Size} - The number of qubits in the system. For example the quantum teleportation protocol has a fixed system size of 3 whereas the Quantum Fourier Transform can scale to any system size.

\textbf{Quantum State} - A column vector of $2^n$ complex numbers representing the probability amplitudes of the $2^n$ states $\ket{0}\dots\ket{2^n-1}$ for a system size $n$.

\textbf{Quantum Gate} - A unitary operation performed on a quantum state.

\textbf{Quantum Circuit} - An ordered list of quantum gates to be applied to the quantum state.

\textbf{Quantum Program} - An ordered list of instructions used to construct a quantum circuit.

\textbf{Quantum Algorithm} - An ordered list of instructions including variables such as the system size.
A instantiation of a quantum algorithm with the system size specified becomes as quantum program.

\textbf{Suitability Measure} - A function that assigns a quantitative measure of how well a proffered solution performs.
A value of $0$ indicates ideal functionality with increasingly higher values of the function indicating ever poorer performance.

\textbf{Search Target} - Every search has a target which is a representation of the area of the search space that contains solutions to the search problem.
For example, in a search for $x$ in a list the search target could be the indices of all occurrences of $x$ in the list and in a search for $x=\sqrt{4}$ the search target would be $x=\pm2$.

\section{Requirements Summaries}
This section contains a summary of the requirements of each of the separate phases of the project.
% A full listing of specific requirements can be found in Appendix \ref{sec:reqs}.

\subsection{Framework}
\subsubsection{Additional Search Engines - Req:ASE}
\label{sec:reqase}
\textbf{The framework shall allow researchers to provide search engines for the system to use.}
Different search techniques are likely to perform better for different problems.
% The framework is also intended for research into the search techniques used for searching for quantum algorithms.
The way in which the framework provides this shall not favour any search technique.
% It is important that the framework shall effectively be research direction independent.

\subsubsection{Additional Suitability Measures - Req:ASM}
\label{sec:reqasm}
\textbf{The framework shall allow researchers to provide suitability measures for the system to use.}
In heuristic search, the terms fitness function and cost functions are typically used to indicate suitability measures.
In general we would seek to maximise ``fitness'' and minimise ``cost'' of any proffered solution.
% A suitability measure is effectively a fitness function.
% However, the term fitness function is associated with the use of evolutionary techniques.
% With the tool intended to be technique independent the term suitability measure shall be used.

It is well known that the suitability measure may have significant impact on the success of a search.
A suitability measure that works well for one problem may not work for another. 
Therefore, the generality of the framework relies heavily on being able to add suitability measures.

\subsubsection{Quantum Algorithm Output - Req:QAO}
\label{sec:reqqao}
\textbf{The solution of a search, a quantum algorithm, shall be presented to the user as a list of instructions.}
An algorithm is a list of instructions parametrised by the system size.
When an instantiation with a specified system size is executed a circuit for that system size is produced.
The solution of a search using the framework is an algorithm as defined in Section \ref{sec:reqdefs}.
This solution shall be provided to the user as a list of instructions in a consistent format.

\subsubsection{Circuit Visualisation  - Req:CV}
\label{sec:reqcv}
\textbf{The system shall provide visualisation of the circuit produced by the solution of the search for a user specified system size.}
To ensure that the output of the search is useful the framework shall provide a representation of the resulting circuit that can be rendered into a circuit diagram.

The circuit visualisations produced use standard depictions of gates and circuits.

\subsubsection{Third Party Software - Req:TPS}
\label{sec:reqtps}
\textbf{The framework shall be able to be embedded in third party software.}
The framework is intended for use by the research community and it is not intended to limit the ways that it can be used.
As a result it is not only important that the framework be able to use third party software, search engines and suitability measures, but is also important for the framework to be available for inclusion in third party software.
% To achieve this knowledge of the internal implementation detail shall not be required.

\subsubsection{Definition of Search Target - Req:DST}
\label{sec:reqdst}
\textbf{The framework shall provide a standardised format for users to define the target of the search.}
% All searches have a target, the shortest path or the minimum value for example.
% The searches that the framework are intended for are those to find a quantum algorithm to produce a circuit to solve a specific problem, to produce an equal superposition or the Quantum Fourier Transform circuit.
The framework needs to provide a standard way of defining the details of a search target.
The standard shall be formalised so it is able to be used and generated by third party software.

\subsubsection{Use of Configuration Files - Req:UCF}
\label{sec:requcf}
\textbf{The customisation of the framework shall be provided through a series of configuration files.}
All third party additions to the framework, search engines and suitability measures, shall be specified using a series of configuration files.
These configuration files shall be well defined and able to be used and produced by third party software.
Files are specified for persistence.

Another option was for configuration to be manually performed each time the framework is used.
This was rejected as the configuration of available search engines and suitability measures are unlikely to change very often.
Maintaining a persistent configuration will also reduce the opportunity for human error during configuration as it will be performed much less frequently.

\subsubsection{Provided Gates and Algorithm Instructions - Req:PGAI}
\label{sec:reqpgai}
\textbf{The framework shall provide implementations of all gates specified in Table \ref{tab:providedgates}.
The framework shall provide algorithm instructions for each of these gates and for the instantiation of the Controlled-U gate with all single qubit gates.}
Table \ref{tab:providedgates} defines most well known quantum gates and indicates the visual representations used in the project.

\begin{table}
 \begin{longtable}{|c c|c c|c c|}
\hline & & & & & \\
$
\Qcircuit @C=1.0em @R=.7em {
&\gate{X}&\qw
}
$
& 
$
\begin{pmatrix}0&&1\\1&&0\end{pmatrix}
$ 
& 
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{Y}&\qw
}
$ 
&
$
\begin{pmatrix}0&&-i\\i&&0\end{pmatrix}
$
&
$
\Qcircuit @C=1.0em @R=.7em {
&\gate{Z}&\qw
}
$
& 
$
\begin{pmatrix}1&&0\\0&&-1\end{pmatrix}
$  \\ & & & & & \\
\hline & & & & & \\
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{H}&\qw
}
$ 
&
$
\begin{pmatrix}\frac{1}{\sqrt{2}}&&\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}&&-\frac{1}{\sqrt{2}}\end{pmatrix}
$ 
&
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{RX}&\qw
}
$ 
&
$
\begin{pmatrix}\cos{\frac{\theta}{2}}&&-i\sin{\frac{\theta}{2}}\\-i\sin{\frac{\theta}{2}}&&\cos{\frac{\theta}{2}}\end{pmatrix}
$
&
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{RY}&\qw
}
$ 
&
$
\begin{pmatrix}\cos{\frac{\theta}{2}}&&-\sin{\frac{\theta}{2}}\\\sin{\frac{\theta}{2}}&&\cos{\frac{\theta}{2}}\end{pmatrix}
$  \\ & & & & & \\
\hline & & & & & \\
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{RZ}&\qw
}
$ 
&
$
\begin{pmatrix}e^{-i\frac{\theta}{2}}&&0\\0&&e^{i\frac{\theta}{2}}\end{pmatrix}
$ 
&
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{V}&\qw
}
$ 
&
$
\begin{pmatrix}1&&0\\0&&i\end{pmatrix}
$
&
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{W}&\qw
}
$ 
&
$
\begin{pmatrix}1&&0\\0&&-i\end{pmatrix}
$  \\ & & & & & \\
\hline & & & & & \\
$
\Qcircuit @C=1.0em @R=.7em {
& \gate{Zero}&\qw
}
$ 
&
$
\begin{pmatrix}1&&1\\0&&0\end{pmatrix}
$ 
&
$
\Qcircuit @C=1.0em @R=.7em {
& \qswap &\qw \\
& \qswap \qwx &\qw
}
$ 
&
$
\begin{pmatrix}a\\b\\c\\d\end{pmatrix} \rightarrow \begin{pmatrix}a\\c\\b\\d\end{pmatrix}
$
&
$
\Qcircuit @C=1.0em @R=.7em {
& \ctrl{1} & \qw \\
& \gate{U} \qwx & \qw
}
$ 
&
$
\begin{pmatrix}I&&0\\0&&U\end{pmatrix}
$  \\
 & & & & & \\
\hline& & & & & \\
& &  $
\Qcircuit @C=1.0em @R=.7em {
& \gate{R(k)} & \qw
}
$ 
&
$
\begin{pmatrix}1&&0\\0&&e^{\frac{2\pi{i}}{2^k}}\end{pmatrix}
$ & & \\
& & & & & \\
\hline
  
 \end{longtable}
\caption{Supported Gates and Definitions}
\label{tab:providedgates}
\end{table}

\subsubsection{Algorithm Control Structures - Req:ACS}
\label{sec:reqacs}
\textbf{The system shall provide an iterate control structure and support nested iterate structures.}
Many algorithms require nested loop structures.
However, na\"{\i}ve nesting is not sufficient.
Each loop construct shall provide a loop variable that must be accessible within the loop including inside any nested loops.
The algorithm produced by Massey\cite{masseythesis} producing the Quantum Fourier Transform circuit is an algorithm that requires such a loop structure.

\subsubsection{Producing Circuits from Algorithms - Req:PCA}
\label{sec:reqpca}
\textbf{The framework shall be able to produce a circuit, for any system size, from a quantum algorithm.}

\subsubsection{Circuit Simulation - Req:CS}
\label{sec:reqcs}
\textbf{The framework shall provide the simulation of a circuit given an initial state.}
Using the gate definitions given in Table \ref{tab:providedgates}, a circuit constructed of the supported gates shall be able to be accurately simulated.
Given an initial state the framework shall be able to give the final state up to the accuracy of floating point arithmetic.

\subsubsection{Step-by-Step State Evolution - Req:SBSSE}
\label{sec:reqsbsse}
\textbf{The framework shall provide a way to perform step-by-step evaluation of a circuit given an initial state.}
To aid researchers in understanding the circuits produced by the preffered algorithms a step-by-step evaluation shall be provided.
Given an initial state and a circuit, the state after the application of each unitary operation, gate, shall be reported so the state evolution can be traced.

This will also provide a debugging mechanism to ensure that all unitary operations are performing the expected opertation on state.

\subsection{Prototype Implementation}
\subsubsection{Sample Search Engine - Req:SSE}
\label{sec:reqsse}
\textbf{The tool shall provide at least one implemented search engine.}
The tool shall provide a basic search engine that will allow researchers interested in the quantum algorithms, rather than the search techniques, to use the tool ``out of the box''.
If time permits more may be provided.

\subsubsection{Sample Suitability Measure - Req:SSM}
\label{sec:reqssm}
\textbf{The tool shall provide at least one implemented suitability measure.}
The tool shall provide a basic suitability measure that will allow researchers interested in the quantum algorithms, rather than the suitability measures, to use the tool ``out of the box''.
The suitability measure provided must allow basic circuits to be produced but is otherwise unspecified.

\subsubsection{Sample Search Targets - Req:SST}
\label{sec:reqsst}
\textbf{The tool shall provide a number of search targets with known outputs.}
To allow search engine and suitability measure researchers to perform simple experiments the tool shall provide a selection of basic search targets.
The search targets included are not specified.

\subsection{Client}
\label{sec:clientguireqs}

\subsubsection{Search Engine Selection - Req:SES}
\label{sec:reqses}
\textbf{The GUI shall provide a user with a selection of search engines to use in a search.}
The GUI shall provide a selection between all search engines registered in the framework.
The most recently selected search engine shall be used by subsequent searches.

\subsubsection{Suitability Measure Selection - Req:SMS}
\label{sec:reqsms}
\textbf{The GUI shall provide a user with a selection of suitability measures to use in a search.}
The GUI shall provide a selection between all suitability measures registered in the framework.
The most recently selected suitability measure shall be used by subsequent searches.

\subsubsection{Search Target Selection - Req:STS}
\label{sec:reqsts}
\textbf{The GUI shall provide a user with a selection of search targets to be used as the search goal.}
The GUI shall provide a selection between all search targets registered in the framework.
The most recently selected search target shall be set for subsequent searches.

\subsubsection{Search Target Creation - Req:STC}
\label{sec:reqstc}
\textbf{The GUI shall provide a way for users to create a new search target without to explicitly writing the configuration file.}
Writing configuration files is quite monotonous and highly error prone.
The GUI shall provide a way to create these configuration files that reduces the error rate.
The way that the GUI provides this is not expected to dramatically decrease the monotony due to the nature of the amount of information required for problems with high numbers of qubits.
The inclusion of such a feature is very important to improve the usability of the system and improve the potential level of use in the research community.

\subsubsection{Search Target Editing - Req:STE}
\label{sec:reqste}
\textbf{The GUI shall provide a way for users to edit the contents of a previously created search target without manual editing the configuration file.}
The size of the data per test case increases exponentially as the number of qubits increases.
The size of the configuration file required to specify a search target will increase linearly on top of this with respect to the number of test cases.
With the size of configuration file increasing in such a dramatic way the risk of error when directly editing the configuration files increases in a similar fashion.
To decrease the risk of errors the GUI shall provide a way to graphically edit the test cases.

\subsubsection{Loading a Search Target From a Previously Defined Configuration File - Req:LSTPDC}
\label{sec:reqlstpdc}
\textbf{The GUI shall provide a way to import a predefined search target from a configuration file.}
One of the intended uses of the GUI is for research into producing quantum algorithms.
It is possible that researchers will want to distribute the search target definitions they create.
This distribution may be to colleagues or simply to other computers for them to continue work.
Either way once a search target is defined and distributed, the use of received search target configuration files should be supported by the GUI.
The GUI shall provide a way for users to import search targets using the respective search target configuration file as long as the configuration file is of the correct format.

\subsubsection{State Visualisation - Req:SV}
\label{sec:reqsv}
\textbf{The GUI shall provide a way to visualise any quantum state.}
A quantum state is defined as a vector of complex numbers.
Depending on size, comparing two or more states can become monotonous.
If the comparison of the two states does not need to be exact, a visual representation of the two states can provide a simpler, and quicker, method for comparison.
To provide such comparison the GUI shall provide a way to visualise a quantum state.

\subsubsection{Reporting the Search Result - Req:RSR}
\label{sec:reqrsr}
\textbf{The GUI shall provide a way to report the search result, a quantum algorithm, to the user.}
The GUI would be of no use to any quantum algorithm researcher if it did not provide the results of a search.
The GUI shall provide the quantum algorithm meeting requirement Req:QAO.

\subsubsection{Graphical Circuit Visualisation - Req:GCV}
\label{sec:reqgcv}
\textbf{Given a quantum algorithm and a system size, the GUI shall produce a visualisation of the resulting circuit.}
Some quantum algorithms produced using the search are likely to be hard to understand in pure algorithm form.
Understanding a circuit is likely to be easier.
To save researcher time in drawing the circuits by hand, the GUI shall provide a visualisation of the circuit for a specified system size.

\subsubsection{Graphical Step-by-Step State Evolution - Req:GSBSSE}
\label{sec:reqgsbsse}
\textbf{The GUI shall provide a way to perform, control and visualise the step-by-step state evolution for an initial state and circuit.}
The framework provides the ability to analyse the evolution of a state with respect to an initial state and a circuit.
The GUI shall provide a way of controlling and reporting this step by step evaluation to the user.

\subsubsection{Tooltips - Req:TT}
\label{sec:reqtt}
\textbf{The GUI shall provide user help through the use of tooltips.}
All elements of the GUI shall be explained through the use of tooltips.
A separate help system would be another option.
Tooltips were chosen over a separate help system as they provide less intrusive help and improved context.
A second consideration is the provision of GUI components by third party.
With a separate help system these GUI components would either have to provide another separate help system or integrate into the main help system.
With tooltips, all help is locally defined and therefore does not require integration making it much simpler when considering potential third party components.

\subsection{General Requirements}
\subsubsection{Portability - Req:POR}
\label{sec:reqpor}
\textbf{The framework, fully implemented tool and the GUI shall be able to be used on a range of Operating Systems.}
The produced software shall be able to be run on:
\begin{itemize}
 \item Windows 7
 \item Linux
\end{itemize}
% 
% \subsubsection{Usability - Req:USE}
% \label{sec:requse}
% \textbf{Using either the fully implemented tool or the GUI a user shall be able to start a search within 30 seconds.}
% Using a predefined search target a user shall be able to initiate a search with a chosen search engine and suitability measure within 30 seconds of starting the software.